---
title: "ADS-506 Team 1 Final Project - EDA"
author: "Jun Clemente"
format: pdf
editor: visual
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(fpp3)
library(gt)
library(skimr)
library(scales)
```

# Dataset

## Quick Summary

```{r import_dataset}
df <- read_csv("../datasets/total_edify_calls_2024-11-03.csv")
# remove spaces from column names
colnames(df) <- gsub(" ", "", colnames(df))
head(df)
```

```{r}
# quick summary of dataframe
summary(df)
```

## Columns with missing values

```{r}
# show columns with missing values
sapply(df, function(x) sum(is.na(x)))
# show rows that have missing values
rows_with_na <- df[!complete.cases(df),]
rows_with_na
# only one row missing values. remove row from dataset
df_clean <- na.omit(df)
df_clean[!complete.cases(df_clean),]
```

```{r}
# check for rows with NA, Inf, or -Inf
rows_with_non_finite <- df_clean %>%
  filter(
    if_any(c(HoldTime, TimeInteracting, WaitTime, WrapUpTime), ~ !is.finite(.))
  )
rows_with_non_finite
```

## Detailed view of data

```{r}
skim(df_clean)
```

### Observation

The dataset has 8 features and 275,655 records.

Two of the features are datetime information.

Two features are categorical. Four features are continuous.

Out of all the records, only one row is missing data.

## Categorical Data

```{r}
df_clean %>%
  count(CommunicationType) %>%
  ggplot(aes(x = CommunicationType, y = n)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = n), vjust = -0.5) +
  labs(
    title = "Frequency of Communication Type by Category",
    x = "Communication Type", 
    y = "Count"
  ) +
  scale_y_continuous(labels = comma) +
  theme_minimal()

df_clean %>%
  count(SubCommunicationType) %>%
  ggplot(aes(x = SubCommunicationType, y = n)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = n), vjust = -0.5) +
  labs(
    title = "Frequency of Sub Communication Type by Category",
    x = "Communication Type", 
    y = "Count"
  ) +
  scale_y_continuous(labels = comma) +
  theme_minimal()

```

### Observations

## Continuous Variables

```{r}
# reshape data to long format
df_long <- df_clean %>%
  pivot_longer(
    cols = c(WaitTime, TimeInteracting, HoldTime, WrapUpTime), 
    names_to = "Variable", values_to = "Value"
  ) %>%
  mutate(Value = Value / 60)

# create box plots
ggplot(df_long, aes(x = "", y = Value)) + 
  geom_boxplot() + 
  facet_wrap(~ Variable, scales = "free_y") + 
  coord_cartesian(ylim = c(0, 5)) +
  labs(
    title = "Distribution of Continuous Variables", 
    x = "Variable", 
    y = "Wait Time (minutes)"
  ) +
  scale_y_continuous( labels = comma ) +
  theme_minimal()

ggplot(df_long, aes(x = Value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ Variable, scales = "free_x") + 
  labs(
    title = "Distsribution",
    x = "Time (minutes)", 
    y = "Frequency"
  ) +
  scale_y_continuous(labels = comma) + 
  scale_x_continuous(labels = comma) + 
  theme_minimal()
```

```{r}
continuous_var <- c("WaitTime", "TimeInteracting", "HoldTime", "WrapUpTime")

skim(df_clean[, continuous_var]) %>%
  select(skim_variable, numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 2
  ) %>%
  cols_label(
    skim_variable = "Variable",
    numeric.mean = "Mean", 
    numeric.sd = "SD", 
    numeric.p0 = "Min", 
    numeric.p25 = "25%", 
    numeric.p50 = "50%", 
    numeric.p75 = "75%", 
    numeric.p100 = "Max"
  ) %>%
  tab_header(
    title = "Statistics for Continuous Variables (Minutes)"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  cols_align(
    align = "center", 
    columns = c(numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100)
  )

```

### Observations

Each of the continuous variables have relatively low means and they also contain extremely high outliers.

# Time Series

## Hourly
```{r}
# aggregate to hourly
df_hourly <- df_clean %>%
  mutate(hour = floor_date(StartTime, "hour")) %>%
  group_by(hour) %>%
  summarise(total_calls = n())

# convert to tsibble
df_hourly_ts <- df_hourly %>%
  as_tsibble(index = hour)

# plot using autoplot
autoplot(df_hourly_ts, total_calls) +
  labs(
    title = "Total Calls", 
    x = "Hour",
    y = "Total Calls"
  ) +
  theme_minimal()

  
```
## Weeklyi
```{r}

df_weekly <- df_clean %>%
  mutate(week = floor_date(StartTime, "week")) %>%
  group_by(week) %>%
  summarise(total_calls = n()) %>%
  ungroup() %>%
  complete(week = seq.Date(min(week), max(week), by = "week"), fill = list(total_calls = 0))
# df_daily <- df_clean %>%
#   mutate(day = floor_date(StartTime, "day")) %>%
#   group_by(day) %>%
#   summarise(total_calls = n()) %>%
#   ungroup() %>%
#   complete(day = seq.Date(min(day), max(day), by = "day"), fill = list(total_calls = 0))

# convert to tsibble

df_weekly_ts <- df_weekly %>%
  as_tsibble(index = week)

# plot chart
df_weekly_ts %>%
  autoplot(total_calls) +
  labs(
    title = "Title"
  ) + 
  theme_minimal()
  



```

