---
title: "ADS-506 Team 1 Final Project - EDA"
author: "Jun Clemente"
format: pdf
editor: visual
---

```{r}
#| message: false
#| warning: false
library(tidyverse)
library(fpp3)
library(gt)
library(skimr)
library(scales)
```

# Dataset

## Quick Summary

```{r import_dataset}
df <- read_csv("../datasets/total_edify_calls_2024-11-03.csv")
# remove spaces from column names
colnames(df) <- gsub(" ", "", colnames(df))
head(df)
```

```{r}
# quick summary of dataframe
summary(df)
```

## Columns with missing values

```{r}
# show columns with missing values
sapply(df, function(x) sum(is.na(x)))
# show rows that have missing values
rows_with_na <- df[!complete.cases(df),]
rows_with_na
# only one row missing values. remove row from dataset
df_clean <- na.omit(df)
df_clean[!complete.cases(df_clean),]
```

```{r}
# check for rows with NA, Inf, or -Inf
rows_with_non_finite <- df_clean %>%
  filter(
    if_any(c(HoldTime, TimeInteracting, WaitTime, WrapUpTime), ~ !is.finite(.))
  )
rows_with_non_finite
```

## Detailed view of data

```{r}
skim(df_clean)
```

### Observation

The dataset has 8 features and 275,655 records.

Two of the features are datetime information.

Two features are categorical. Four features are continuous.

Out of all the records, only one row is missing data.

## Categorical Data

```{r}
df_clean %>%
  count(CommunicationType) %>%
  ggplot(aes(x = CommunicationType, y = n)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = n), vjust = -0.5) +
  labs(
    title = "Frequency of Communication Type by Category",
    x = "Communication Type", 
    y = "Count"
  ) +
  scale_y_continuous(labels = comma) +
  theme_minimal()

df_clean %>%
  count(SubCommunicationType) %>%
  ggplot(aes(x = SubCommunicationType, y = n)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = n), vjust = -0.5) +
  labs(
    title = "Frequency of Sub Communication Type by Category",
    x = "Communication Type", 
    y = "Count"
  ) +
  scale_y_continuous(labels = comma) +
  theme_minimal()

```

### Observations

## Continuous Variables

```{r}
# reshape data to long format
df_long <- df_clean %>%
  pivot_longer(
    cols = c(WaitTime, TimeInteracting, HoldTime, WrapUpTime), 
    names_to = "Variable", values_to = "Value"
  ) %>%
  mutate(Value = Value / 60)

# create box plots
ggplot(df_long, aes(x = "", y = Value)) + 
  geom_boxplot() + 
  facet_wrap(~ Variable, scales = "free_y") + 
  coord_cartesian(ylim = c(0, 5)) +
  labs(
    title = "Distribution of Continuous Variables", 
    x = "Variable", 
    y = "Wait Time (minutes)"
  ) +
  scale_y_continuous( labels = comma ) +
  theme_minimal()

ggplot(df_long, aes(x = Value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ Variable, scales = "free_x") + 
  labs(
    title = "Distsribution",
    x = "Time (minutes)", 
    y = "Frequency"
  ) +
  scale_y_continuous(labels = comma) + 
  scale_x_continuous(labels = comma) + 
  theme_minimal()
```

```{r}
continuous_var <- c("WaitTime", "TimeInteracting", "HoldTime", "WrapUpTime")

skim(df_clean[, continuous_var]) %>%
  select(skim_variable, numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100) %>%
  gt() %>%
  fmt_number(
    columns = everything(),
    decimals = 2
  ) %>%
  cols_label(
    skim_variable = "Variable",
    numeric.mean = "Mean", 
    numeric.sd = "SD", 
    numeric.p0 = "Min", 
    numeric.p25 = "25%", 
    numeric.p50 = "50%", 
    numeric.p75 = "75%", 
    numeric.p100 = "Max"
  ) %>%
  tab_header(
    title = "Statistics for Continuous Variables (Minutes)"
  ) %>%
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(everything())
  ) %>%
  cols_align(
    align = "center", 
    columns = c(numeric.mean, numeric.sd, numeric.p0, numeric.p25, numeric.p50, numeric.p75, numeric.p100)
  )

```

### Observations

Each of the continuous variables have relatively low means and they also contain extremely high outliers.

# Time Series

## Hourly

```{r}
# aggregate to hourly
df_hourly <- df_clean %>%
  mutate(hour = floor_date(StartTime, "hour")) %>%
  group_by(hour) %>%
  summarise(total_calls = n())

# convert to tsibble
df_hourly_ts <- df_hourly %>%
  as_tsibble(index = hour)

# plot using autoplot
autoplot(df_hourly_ts, total_calls) +
  labs(
    title = "Total Calls", 
    x = "Hour",
    y = "Total Calls"
  ) +
  theme_minimal()
```

## Weekly - Total Call Volume

```{r}
df_weekly <- df_clean %>%
  mutate(week = as.Date(StartTime, "week")) %>%  # Round StartTime to the beginning of the week
  group_by(week) %>%
  summarise(total_calls = n()) %>%  # Count the number of rows (calls) per week
  ungroup() %>%
  # Fill in missing weeks with 0 calls
  complete(week = seq.Date(min(week), max(week), by = "week"), fill = list(total_calls = 0))  

df_weekly_ts <- df_weekly %>%
  as_tsibble(index = week)

# plot chart
df_weekly_ts %>%
  autoplot(total_calls) +
  labs(
    title = "Weekly Call Volumes (Mar 2022 - Oct 2024)",
    y = "Total Calls",
    x = "Date"
  ) + 
  theme_minimal()

# decomp of weekly total call volume
decomp_calls <- df_weekly_ts %>%
  fill_gaps(total_calls = 0) %>%
  model(stl = STL(total_calls ~ season(window = "periodic")))

# extract and view decomp components
components_calls <- decomp_calls %>%
  components()

# plot decomp
components_calls %>%
  autoplot() +
  labs(
    title = "STL Decomposition of Weekly Call Volumnes", 
    y = "Total Calls", 
    x = "Date"
  ) +
  theme_minimal()
```

### Observations

Time Series dataset is missing some data that requires filling in. To create these charts, missing weeks were arbitrarily filled in with "0".

The Remainder chart does not appear to be random. This would need to be explored further to determine if there are any uncaptured season trends, or autocorrelations.

## Weekly - Average of WaitTime

```{r}
df_weekly_wait <- df_clean %>%
  mutate(week = as.Date(StartTime , "week")) %>%
  group_by(week) %>%
  summarise(avg_wait_time = mean(WaitTime, na.rm = TRUE)) %>%
  ungroup() %>%
  # Fill in missing weeks
  complete(week = seq.Date(min(week), max(week), by = "week"), fill = list(avg_wait_time = 0))

# convert to tsibble
df_weekly_wait_ts <- df_weekly_wait %>%
  as_tsibble(index = week)

# plot chart
df_weekly_wait_ts %>%
  autoplot(avg_wait_time) +
  labs(
    title = "Weekly Average Wait Time (Mar 2022 - Oct 2024)", 
    y = "Average Wait Time (min)",
    x = "Date"
  ) +
  theme_minimal()
```

```{r}
decomp_wait <- df_weekly_wait_ts %>%
  fill_gaps() %>%
  mutate(avg_wait_time = if_else(is.na(avg_wait_time), mean(avg_wait_time, na.rm = TRUE), avg_wait_time)) %>%
  model(stl = STL(avg_wait_time ~ season(window = "periodic")))

# extract and view decomp components
components_wait <- decomp_wait %>%
  components()

# plot decomp
components_wait %>%
  autoplot() + 
  labs(
    title = "STL Decomposition of Weekly Average Wait Time", 
    y = "Average Wait Time", 
    x = "Date"
  ) +
  theme_minimal()
```

### Observations

Based on average wait time, this does not have any seasonality or cyclic elements. Some weeks were missing data and therefore arbitrarily imputed with the mean wait time.
